True versus apparent shapes of astrophysical bow shocks

### 2017-03-27 papers/Paper1/sec-quadrics-observations.tex
Describe star rating system

The most extensive sample of stellar bow shock nebulae to date is a
catalog of 709 arcs \citep{Kobulnicky:2016a} detected in mid-infrared
surveys of the Galactic Plane by the \textit{Spitzer Space Telescope}
(\textit{SST}, \citealp{Werner:2004a}) and \textit{Wide-field Infrared
  Survey Explorer} (\textit{WISE}, \citealp{Wright:2010a}).  These
sources are believed to be powered by the winds of early-type stars,
which are either moving supersonically through the interstellar medium
(runaway stars), or are interacting with a local bulk flow, such as
the champagne flow from a nearby \hii{} region.

@@

We developed a methodology for automatically tracing the arcs as follows:
\begin{enumerate}[1.]
\item Calculate arrays of celestial coordinates, \(C\), for each pixel
  of the image.
\item Using the central source coordinates, \(C_0\) and nominal
  bowshock radius, \(R_0\) from \citet{Kobulnicky:2016a}, construct a
  pixel mask that includes only those pixels with separations
  \(|C - C_0|\) from the source that lie between \(0.5 R_0\) and
  \(3 R_0\).  This mask will be used for all subsequent operations,
  which serves to help avoid confusion from the star itself and other
  bright sources in the field of view.
\item Define a ``step-back'' point, \(C_1\), which is located at a
  separation \(2 R_0\) from the source, but in the opposite direction
  from the apex of the bow shock. That is, along a position angle
  180\degr{} from the nominal position angle, \(\text{PA}_0\), of
  the bow shock axis.
\item Looping over a grid of 50 position angles, \(\text{PA}_k\), within
  \(\pm 60\degr\) of \(\text{PA}_0\), estimate the location of the
  arc along each PA from the step-back point via two different methods:
  \begin{enumerate}[(a)]
  \item The pixel with the maximum brightness, with coordinates
    \(C_{k,\text{peak}}\) (red dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \item The mean brightness-weighted separation from \(C_1\), with
    coordinates \(C_{k,\text{mean}}\) (light blue dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \end{enumerate}
  For each \(\text{PA}_k\) in the grid, the calculation is performed
  over only those pixels that satisfy
  \(|\text{PA}(C, C_1) - \text{PA}_k| < \frac12 \delta\theta\), where
  \(\delta\theta = 120/50 = 2.4\degr\), which defines a thin radial wedge from
  \(C_1\).
\item For each of the points \(C_{k,\text{peak}}\) and \(C_{k,\text{mean}}\), determine the radial separation from the central source, \(R_k = |C_k - C_0|\) and the angle from the bow shock axis about the central source \(\theta_k = \text{PA}(C_k, C_0) - \text{PA}_0\).  These are plotted in the upper left panels of  Figure~\ref{fig:mipsgal-examples}.

\end{enumerate}

>>>>

The most extensive observational sample of stellar bow shock nebulae
to date is a catalog of 709 arcs \citep{Kobulnicky:2016a} detected in
mid-infrared surveys of the Galactic Plane by the \textit{Spitzer
  Space Telescope} (\textit{SST}, \citealp{Werner:2004a}) and
\textit{Wide-field Infrared Survey Explorer} (\textit{WISE},
\citealp{Wright:2010a}).  These sources are believed to be powered by
the winds of early-type stars, which are either moving supersonically
through the interstellar medium (runaway stars), or are interacting
with a local bulk flow, such as the champagne flow from a nearby
\hii{} region.

@@

We developed a methodology for automatically tracing the arcs as follows:
\begin{enumerate}[1.]
\item Calculate arrays of celestial coordinates, \(C\), for each pixel
  of the image.
\item Using the central source coordinates, \(C_0\) and nominal
  bowshock radius, \(R_0\) from \citet{Kobulnicky:2016a}, construct a
  pixel mask that includes only those pixels with separations from the
  source that satisfy \(\frac12 R_0 \le |C - C_0| \le 3 R_0\).  This mask
  will be used for all subsequent operations, which serves to help
  avoid confusion from the star itself and other bright sources in the
  field of view.
\item Define a ``step-back'' point, \(C_1\), which is located at a
  separation \(2 R_0\) from the source, but in the opposite direction
  from the apex of the bow shock. That is, along a position angle
  180\degr{} from the nominal position angle, \(\text{PA}_0\), of
  the bow shock axis.
\item Looping over a grid of 50 position angles, \(\text{PA}_k\), within
  \(\pm 60\degr\) of \(\text{PA}_0\), estimate the location of the
  arc along each PA from the step-back point via two different methods:
  \begin{enumerate}[(a)]
  \item The pixel with the maximum brightness, with coordinates
    \(C_{k,\text{peak}}\) (red dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \item The mean brightness-weighted separation from \(C_1\), with
    coordinates \(C_{k,\text{mean}}\) (light blue dots in
    Fig.~\ref{fig:mipsgal-examples}).
  \end{enumerate}
  For each \(\text{PA}_k\) in the grid, the calculation is performed
  over only those pixels that satisfy
  \(|\text{PA}(C, C_1) - \text{PA}_k| < \frac12 \delta\theta\), where
  \(\delta\theta = 120/50 = 2.4\degr\), which defines a thin radial wedge from
  \(C_1\).  The results are shown as red and blue dots superimposed
  on the images in Figure~\ref{fig:mipsgal-examples}. Each of the two
  methods, ``peak'' and ``mean'', works better in some objects and
  worse in others (according to the subjective judgment of
  ``correctly'' tracing the bow shock shape).  We therefore take the
  average by amalgamating all the \(C_{k,\text{peak}}\) and
  \(C_{k,\text{mean}}\) points into a single set, \(C_{k}\), for
  subsequent analysis.
\item For each of the points \(C_{k}\), determine the radial
  separation from the central source, \(R_k = |C_k - C_0|\) and the
  angle from the bow shock axis about the central source
  \(\theta_k = \text{PA}(C_k, C_0) - \text{PA}_0\).  These are plotted in
  the upper left panels of Figure~\ref{fig:mipsgal-examples}.
\item Make our own estimate of the axial size, \(R_0\), of the bow
  shock by calculating the mean and standard deviation of \(R_k\) over
  all points \(C_k\) with \(|\theta_k| \le 10\degr\).  Note that this is
  distinct from the nominal value of \(R_0\) given in the
  \citet{Kobulnicky:2016a} catalog, which was ``measured by eye''.
\item Estimate the radius of curvature, \(R_c\), by fitting a circle
  to all those points within \(\pm 45\degr\) of the nominal axis
  (\(|\theta_k| < 45\degr\)), but after excluding any point with
  \(R_k < \frac12 R_m\) or \(R_k > 2 R_m\), where \(R_m\) is the median
  \(R_k\) for \(|\theta_k| < 45\degr\).
\item Determine two separate estimates, \(R_{90+}\) and \(R_{90-}\),
  of the perpendicular radius, \(R_{90}\), by taking the mean and
  standard deviation of \(R_k\) over all points \(C_k\) with
  \(|\theta_k - 90\degr| \le 10\degr\) for \(R_{90+}\), and with
  \(|\theta_k + 90\degr| \le 10\degr\) for \(R_{90-}\).
\end{enumerate}

After these automatic steps, we subjectively evaluate the results by giving a star rating to each source:

\paragraph*{0 stars} The fitting algorithm failed for some reason.

\paragraph*{1 star} The fit was formally successful, but the results
for \(R_c\) or \(R_{90}\) are far removed from what a human would
predict by looking at the image.  For example, in the smallest
bowshocks, which are only marginally resolved by Spitzer's 6\arcsec{}
beam, the dispersion in \(R_k\) can be a significant fraction of
\(R_0\), in which case our algorithm tends to erroneously favor
\(R_c < R_0\).

\paragraph*{2 stars} The fit results are not totally outlandish, but
nonetheless some problem is apparent that casts doubt on the
reliability of the results.  For example, a double-shell structure to
the bow shock that leads to large differences between the ``peak'' and
``mean'' methods, or point sources near to the bow shock that
interfere with the tracing procedure.

\paragraph*{3 stars} A good fit, but where the dispersion in \(R_k\)
and/or the asymmetry in the bow shock reduces the precision in the
determination of \(R_c\) and \(R_{90}\), giving subjectively estimated
uncertainties around the 20\% level.  An example of a 3-star fit is
shown in Figure~\ref{fig:mipsgal-examples}a.

\paragraph*{4 stars} A high quality fit, with subjectively estimated
uncertainties in \(R_c\) and \(R_{90}\) around the 10\% level. An
example of a 4-star fit is shown in
Figure~\ref{fig:mipsgal-examples}b.

\paragraph*{5 stars} The highest-quality fit, usually corresponding to
large, sharply defined bow shocks, whose shape is determined with high
precision. An example of a 5-star fit is shown in
Figure~\ref{fig:mipsgal-examples}c.

\bigskip
Only sources with a 3-star or higher rating are used in the following analysis, which amount to approximately half (227 out of 471) of all the MIPSGAL arc sources.
